---
title: "Class 12: RNAseq Galaxy"
author: "Angela Bartolo PID: A15932451"
format: pdf
---

## Section 1. Proportion of G/G in a population.

Downloaded a CSV file from Ensemble 

Here we read the CSV file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl) *100
```

Now let's look at a different population- the GBR population.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(gbr)
```

Find proportion of G|G 

```{r}
table(gbr$Genotype..forward.strand.) / nrow(gbr) *100
```

This varient that is associated with childhood asthma is more frequent in the GBR population than the MXL populaiton. 

Let's dig further


## Section 4: Population Scale Analysis

One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale. 

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. 

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

Sample size of each genotype. 

```{r}
nrow(expr)
table(expr$geno)
```

Median expression level for each of these genotypes 

A/A
```{r}
median( expr[expr$geno == "A/A", "exp"] )
```

A/G
```{r}
median( expr[expr$geno == "A/G", "exp"] )
```

G/G
```{r}
median( expr[expr$geno == "G/G", "exp"] )
```

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?


Lets make a boxplot

```{r}
library(ggplot2)
```

```{r}
ggplot(expr) +
  aes(geno, exp, fill=geno) +
  geom_boxplot(notch=T) 
```

The expression levels between A/A and G/G differ a lot so the SNP effects the expression level of ORMDL3. 